% Appendix for UAV-HFACS Paper
% Complete UAV Enhancement Patterns Documentation
% This file contains supplementary material for the research paper:
% "UAV Accident Forensics using HFACS and Large Language Models"

\appendixtitles{yes}
\appendixstart
\appendix
\section{UAV Enhancement Patterns}
\label{appendix:uav_patterns}
This appendix contains the complete set of UAV enhancement patterns extracted and validated in this study, organized according to the four HFACS~8.0 hierarchical levels. Each entry includes its definition, data source, key indicators, and mapping to the HFACS categories.

% Custom command for pattern entries
\newcommand{\patternentry}[5]{%
  \noindent\textbf{#1}.\\
  \textit{Definition:} #2\\
  \textit{Source:} #3\\
  \textit{Indicators:} #4\\
  \textit{HFACS Mapping:} #5\par\smallskip
}

\subsection*{Level 4 --- Organizational Influences (6 patterns)}

\patternentry{Part\_107\_Compliance}
{Organizational challenges in maintaining compliance with FAA Part 107 small UAS regulations.}
{Regulatory analysis; expert validation.}
{Policy gaps; training deficiencies; compliance monitoring failures.}
{OP000 --- Organizational Policy/Process Issues.}

\patternentry{LAANC\_Authorization}
{Organizational processes for Low Altitude Authorization and Notification Capability.}
{ASRS incident analysis; regulatory framework review.}
{Authorization delays; process confusion; system integration issues.}
{OP000 --- Organizational Policy/Process Issues.}

\patternentry{VLOS\_Requirements}
{Organizational management of Visual Line of Sight operational requirements.}
{Literature review; incident analysis.}
{Policy interpretation; observer coordination; range limitations.}
{OP000 --- Organizational Policy/Process Issues.}

\patternentry{BVLOS\_Operations}
{Organizational challenges in Beyond Visual Line of Sight operations.}
{Expert validation; regulatory analysis.}
{Waiver management; risk assessment; technology integration.}
{OR000 --- Resource Management Problems.}

\patternentry{Battery\_Constraints}
{Organizational resource management for power system limitations.}
{Technical documentation analysis; incident reports.}
{Fleet management; maintenance scheduling; replacement policies.}
{OR000 --- Resource Management Problems.}

\patternentry{Energy\_Management}
{Organizational strategies for UAV energy and endurance management.}
{Literature review; expert validation.}
{Mission planning; battery policies; emergency procedures.}
{OR000 --- Resource Management Problems.}

\subsection*{Level 3 --- Unsafe Supervision (5 patterns)}

\patternentry{GCS\_Interface\_Complexity}
{Supervisory challenges in managing Ground Control Station interface complexity.}
{Expert validation; incident analysis.}
{Training adequacy; interface standardization; workload management.}
{SI000 --- Inadequate Supervision.}

\patternentry{Delayed\_Feedback\_Systems}
{Supervisory management of delayed feedback in remote operations.}
{Literature review; technical analysis.}
{Latency awareness; compensation training; system limitations.}
{SI000 --- Inadequate Supervision.}

\patternentry{LAANC\_Integration}
{Supervisory oversight of LAANC system integration and usage.}
{Regulatory analysis; expert validation.}
{Process supervision; authorization tracking; compliance monitoring.}
{SP000 --- Planned Inappropriate Operations.}

\patternentry{Airspace\_Authorization}
{Supervisory management of airspace authorization processes.}
{ASRS analysis; regulatory framework review.}
{Authorization verification; boundary monitoring; violation prevention.}
{SP000 --- Planned Inappropriate Operations.}

\patternentry{Emergency\_Procedures}
{Supervisory oversight of UAV-specific emergency procedures.}
{Expert validation; incident analysis.}
{Procedure adequacy; training verification; response coordination.}
{SI000 --- Inadequate Supervision.}

\subsection*{Level 2 --- Preconditions for Unsafe Acts (11 patterns)}

\patternentry{C2\_Link\_Reliability}
{Command and Control link reliability as precondition for safe operations.}
{ASRS incident analysis; technical documentation.}
{Signal strength; interference levels; backup systems.}
{PE200 --- Technological Environment.}

\patternentry{Telemetry\_Accuracy}
{Accuracy and timeliness of telemetry data transmission.}
{Technical analysis; expert validation.}
{Data integrity; transmission delays; sensor calibration.}
{PE200 --- Technological Environment.}

\patternentry{Range\_Limitations}
{Communication and control range limitations affecting operations.}
{Literature review; incident analysis.}
{Distance constraints; terrain effects; equipment limitations.}
{PE200 --- Technological Environment.}

\patternentry{Signal\_Degradation}
{Progressive degradation of communication signals during operations.}
{ASRS analysis; technical documentation.}
{Quality metrics; environmental factors; equipment aging.}
{PE200 --- Technological Environment.}

\patternentry{Mode\_Confusion}
{Confusion between different flight modes and automation levels.}
{Expert validation; incident analysis.}
{Mode awareness; transition clarity; interface design.}
{PC200 --- Mental-Awareness.}

\patternentry{Automation\_Dependency}
{Over-reliance on automated systems affecting manual skills.}
{Literature review; expert validation.}
{Skill degradation; system trust; manual proficiency.}
{PC200 --- Mental-Awareness.}

\patternentry{Weather\_Sensitivity}
{UAV sensitivity to weather conditions affecting operations.}
{ASRS analysis; environmental data.}
{Wind limits; visibility requirements; precipitation effects.}
{PE100 --- Physical Environment.}

\patternentry{Spatial\_Disorientation}
{Loss of spatial orientation due to remote operation characteristics.}
{Expert validation; incident analysis.}
{Reference loss; perspective confusion; depth perception.}
{PC200 --- Mental-Awareness.}

\patternentry{Visual\_Limitations}
{Limited visual references and environmental cues in remote operations.}
{Literature review; expert validation.}
{Camera limitations; field of view; resolution constraints.}
{PC200 --- Mental-Awareness.}

\patternentry{Battery\_Limits}
{Battery capacity and performance limitations affecting operations.}
{Technical documentation; incident analysis.}
{Capacity degradation; temperature effects; age factors.}
{PE200 --- Technological Environment.}

\patternentry{Automation\_Degradation}
{Degradation of automated system performance over time.}
{Expert validation; technical analysis.}
{System aging; calibration drift; software issues.}
{PE200 --- Technological Environment.}

\subsection*{Level 1 --- Unsafe Acts (4 patterns)}

\patternentry{Autopilot\_Interactions}
{Errors in interaction with autopilot systems and mode management.}
{ASRS incident analysis; expert validation.}
{Mode selection errors; parameter mistakes; system conflicts.}
{AE100 --- Skill-Based Errors.}

\patternentry{Flight\_Mode\_Switching}
{Errors during transitions between different flight modes.}
{Expert validation; incident analysis.}
{Transition timing; mode awareness; control confusion.}
{AE100 --- Skill-Based Errors.}

\patternentry{Limited\_Visual\_References}
{Errors due to limited visual references in remote operations.}
{Literature review; expert validation.}
{Depth misjudgment; obstacle detection; spatial errors.}
{AE200 --- Perceptual Errors.}

\patternentry{Delayed\_Feedback}
{Errors caused by delayed feedback from remote systems.}
{Expert validation; technical analysis.}
{Overcorrection; timing errors; response delays.}
{AE200 --- Perceptual Errors.}

\subsection*{Theoretical Categories (5 categories)}

\noindent\textbf{Remote Operation Challenges}. \textit{Patterns:} Spatial Disorientation; Visual Limitations; Limited Visual References; Delayed Feedback; GCS Interface Complexity; Delayed Feedback Systems.\par\smallskip

\noindent\textbf{Automation Complexity}. \textit{Patterns:} Mode Confusion; Automation Dependency; Autopilot Interactions; Flight Mode Switching; Automation Degradation.\par\smallskip

\noindent\textbf{Communication Dependencies}. \textit{Patterns:} C2 Link Reliability; Telemetry Accuracy; Range Limitations; Signal Degradation.\par\smallskip

\noindent\textbf{Regulatory Considerations}. \textit{Patterns:} Part 107 Compliance; LAANC Authorization; VLOS Requirements; BVLOS Operations; LAANC Integration; Airspace Authorization.\par\smallskip

\noindent\textbf{Power System Limitations}. \textit{Patterns:} Battery Constraints; Energy Management; Battery Limits; Weather Sensitivity.\par\smallskip

\section{System Access and Usage}
\label{appendix:system_access}

% Additional appendix sections can be added here as needed
